### 2021-09-07 paper/localpsf.tex
added .tex files
writing well-posedness proof

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,r.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}). Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}). Proofs for results in this section are given in Appendix \ref{app:proofs}.

In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.  Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to perform two new Poisson PDE solves when a new point is added.
%, rather than $2(r+1)$ new PDE solves.

It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.

\begin{thm}[Poisson weighting functions are optimal]
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, and let $\gamma_y$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $\gamma_y$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{\gamma_y \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma_y}_{L^2(\Omega)}^2 \\
\text{such that} &\quad \gamma_y(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}

@@

		Initialize constant function $C(x) = 1$

		Compute Lebesgue measure of the domain, $|\Omega|$

		\tcp{Compute Poisson impulse responses $\{\psi_i\}_{i=1}^r$ and Poisson squared impulse responses $\{\theta_i\}_{i=1}^r$}

		\For{$i=1,2, \dots, r$}{
			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.

			Compute $\psi_i$ by solving the Neumann Poisson problem
			\begin{equation}
			\label{eq:psi_eq}
			\begin{cases}
			\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
			\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
			\end{cases}
			\end{equation}
			with condition $\int_\Omega \psi_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.

			Compute $\theta_i$ by solving the Neumann Poisson problem
			\begin{equation}
			\label{eq:theta_eq}
						\begin{cases}
			\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
			\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
			\end{cases}
			\end{equation}
			with condition $\int_\Omega \theta_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.

@@

		\tcp{Compute weighting functions $\{w_i\}_{i=1}^r$}

@@

Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite. Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\AmatPcSymPlus$ contains the null-space of $\AmatPcSym$. This is important when the operator $\mathcal{A}$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero. Hessians in ill-posed distributed parameter inverse problems typically have both of these properties. Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\AmatPcSymPlus$ to be close to the corresponding moderate and large positive eigenvalues of $\AmatPcSym$. I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\AmatPcSym$ as little as possible. Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$. There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad). This is shown in Figure FIG. We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.

@@

\begin{itemize}
	\item computational complexity in terms of PDE solves:
	\item cost of global low rank approximation (data scalability)
	\item cost to build approximation
	\item cost without compression to do a solve
	\item cost with compression to do a solve
	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log N)$ PDE solves, C is big, k is big. Here $10$
	\item HODLR: C is smaller, but k is bigger, vs. H1 peeling process
	\item H-matrix operations cost for our method solve
	\item irregular grid
\end{itemize}

@@

\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
	\label{lem:T_is_linear}
	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with a bounded extension operator, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$. Consider the optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense.
	The following statements hold.
	\begin{enumerate}
		\item[(a)] There exists a unique solution to \eqref{eq:generic_optimization_initial}.
		\item[(b)] 	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
		\begin{equation}
		\label{eq:opt_soln_operator}
		\mathcal{T}(f) := \gamma^*,
		\end{equation}
		where $\gamma^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
	\end{enumerate}
\end{lem}
\begin{proof}
	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is embedded in $C^0(\Omega)$. WORK IN PROGRESS
\end{proof}

@@

\section*{Acknowledgments}
We thank J.J. Alger and Longfei Gao for helpful discussions.

>>>>

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,r.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}). Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).

In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently compute the functions $w_i$. Briefly:
\begin{enumerate}
\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
\begin{equation}
\label{eq:psi_eq}
\begin{cases}
\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \psi_i(x) dx = 0$. Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
\begin{equation}
\label{eq:theta_eq}
\begin{cases}
\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \theta_i(x) dx = 0$.
%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source. We call the solution to this first Poisson problem $\psi_i$.
%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source. We call the solution to this second Poisson problem $\theta_i$.
\item We form $w_i$ as a linear combination of the functions $\theta_i$, and the constant function $C(x)=1$. Details for this linear combination are given in Algorithm \ref{alg:weighting_functions_incremental}. 
\end{enumerate}
In Theorem \ref{thm:wi_alg_is_correct}, we prove that the functions $w_i$ produced by this process indeed solve optimization problem \eqref{eq:wi_optimization_problem}.

Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---by storing the functions $\psi_i$ and $\theta_i$, we only need to perform two new Poisson PDE solves when a new point is added.
%, rather than $2(r+1)$ new PDE solves.

It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension of the domain is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-posed in three spatial dimensions or fewer. In four or higher spatial dimensions, one could construct appropriate weighting functions using radial basis functions, though we do not discuss this here.

\begin{thm}[Poisson weighting functions are optimal]
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, $d \in \{1,2,3\}$, and let $\gamma_y$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $\gamma_y$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{\gamma_y \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma_y}_{L^2(\Omega)}^2 \\
\text{such that} &\quad \gamma_y(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}

@@

%		Initialize constant function $C(x) = 1$

%		Compute Lebesgue measure of the domain, $|\Omega|$

%		\tcp{Compute Poisson impulse responses $\{\psi_i\}_{i=1}^r$ and Poisson squared impulse responses $\{\theta_i\}_{i=1}^r$}

		\For{$i=1,2, \dots, r$}{
%			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.

			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\psi_i$.

%			Compute $\psi_i$ by solving the Neumann Poisson problem
%			\begin{equation}
%			\label{eq:psi_eq}
%			\begin{cases}
%			\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
%			\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
%			\end{cases}
%			\end{equation}
%			with condition $\int_\Omega \psi_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.

			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\theta_i$.

%			Compute $\theta_i$ by solving the Neumann Poisson problem
%			\begin{equation}
%			\label{eq:theta_eq}
%						\begin{cases}
%			\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
%			\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
%			\end{cases}
%			\end{equation}
%			with condition $\int_\Omega \theta_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.

@@

%		\tcp{Compute weighting functions $\{w_i\}_{i=1}^r$}

@@

Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite. Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\AmatPcSymPlus$ contains the null-space of $\AmatPcSym$. This is important when the operator $\mathcal{A}$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero. Hessians in ill-posed distributed parameter inverse problems typically have both of these properties. Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\AmatPcSymPlus$ to be close to the corresponding moderate and large positive eigenvalues of $\AmatPcSym$. I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\AmatPcSym$ as little as possible. Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$. There is a tradeoff: if we choose a larger value for $s(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad). This is shown in Figure FIG. We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.

@@

c\begin{itemize}
	\item computational complexity in terms of PDE solves:
	\item cost of global low rank approximation (data scalability)
	\item cost to build approximation
	\item cost without compression to do a solve
	\item cost with compression to do a solve
	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log N)$ PDE solves, C is big, k is big. Here $10$
	\item HODLR: C is smaller, but k is bigger, vs. H1 peeling process
	\item H-matrix operations cost for our method solve
	\item irregular grid
\end{itemize}

@@

\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
	\label{lem:T_is_linear}
	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with a bounded extension operator, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$. Consider the optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense.
	The following statements hold.
	\begin{enumerate}
		\item[(a)] There exists a unique solution to \eqref{eq:generic_optimization_initial}.
		\item[(b)] 	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
		\begin{equation}
		\label{eq:opt_soln_operator}
		\mathcal{T}(f) := \gamma^*,
		\end{equation}
		where $\gamma^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
	\end{enumerate}
\end{lem}
\begin{proof}
	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is continuously imbedded in $C^0(\Omega)$, so pointwise evaluation $u \mapsto u(x_i)$ of functions at a point $x_i \in \Omega$ is a continuous linear functional on $H^2_N$. The set 
	\begin{equation}
	\label{eq:inverse_image_of_point}
	\{u \in H^2_N : u(x_i)=f_i\}
	\end{equation}
	is therefore a closed affine set in $H^2_N$, because it is the inverse image of a point under a continuous linear functional. Hence the feasible set for optimization problem \eqref{eq:generic_optimization_initial},
	\begin{equation}
	\label{eq:wi_feasible_set}
	X := \{u \in H^2_N : u(x_i)=f_i~\text{for}~ i=1,\dots,r\},
	\end{equation}
	is a closed affine set in $H^2_N$, because it is the intersection of the sets in \eqref{eq:inverse_image_of_point} for $i=1,\dots,r$. The feasible set $X$ is convex because it is affine, and it is non-empty because interpolating the data pairs $\{(x_i, f_i)\}_{i=1}^r$ using any smooth interpolation method yields an element of $X$. The objective function, $\gamma \mapsto \frac{1}{2}||\Delta \gamma||^2$, is convex because it

	We now show that the objective function is strongly convex on the tangent space to $X$,
	\begin{equation*}
	X_0 := \{u \in H^2_N : u(x_i)=0~\text{for}~ i=1,\dots,r\}.
	\end{equation*}

@@

\section*{Acknowledgments}
We thank J.J. Alger, Longfei Gao, and Rami Nammour for helpful discussions.
